Avoiding the Global Sort: \break A Faster Contour Tree Algorithm

### 2017-05-06 paint-contour-tree.tex
typos fixed

We detail the basic definitions about contour trees, following the terminology of Chapter 6 of Carr's thesis \cite{c-tmi-04}.
All our assumptions and definitions
are standard for results in this area, though there is some variability in notation.
The input is a continuous piecewise-linear function $f:\MM \to \RR$, where $\MM$ is a simply connected and fully triangulated simplicial complex of a topological ball in $\RR^d$,
except for specially designated \emph{boundary facets}. We assume that all pairs of vertices from $\MM$ have distinct function values, except for pairs from the same boundary facet,
and that $f$ is only explicitly defined on the vertices, and all other values are obtained by linear interpolation.

@@

We now give a detailed description of the main algorithm.
The primary challenge is to use the data structures to rapidly find
ripe vertices.
An simple example of the execution of the algorithm is given in \Sec{example}.

@@

\section{Divide and conquer through contour surgery} \label{sec:surgery}
Now that we have an efficient join tree algorithm, we now describe how to 
reduce computing the contour tree to computing the join tree, as discussed in \Sec{approach}.
Specifically, in this section we first describe our basic cutting tool, which we call contour surgery.
In \Sec{rain} we then describe how we use this tool to cut the surface into so called extremum dominant pieces 
where computing the contour tree reduces to computing the join tree, and this equivalence is proven in \Sec{extreme}. 
The pieces we create are technically not simplicial complexes,
in that each piece may have facets that are not simplices. We may
create facets of larger size than the dimension. These can
easily be completed into simplicial complexes, though as we do not
need this, for ease of exposition, we do not explicitly
perform this step in our algorithm.

>>>>

We detail the basic definitions about contour trees, following the terminology of Chapter 6 of Carr's thesis \cite{c-tmi-04}.
All our assumptions and definitions
are standard for results in this area, though there is some variability in notation.
The input is a continuous piecewise-linear function $f:\MM \to \RR$, where $\MM$ is a fully triangulated simplicial complex of a topological ball in $\RR^d$,
except for specially designated \emph{boundary facets}. We assume that all pairs of vertices from $\MM$ have distinct function values, except for pairs from the same boundary facet,
and that $f$ is only explicitly defined on the vertices, and all other values are obtained by linear interpolation.

@@

We now give a detailed description of the main algorithm.
The primary challenge is to use the data structures to rapidly find
ripe vertices.
A simple example of the algorithm execution is given in \Sec{example}.

@@

\section{Divide and conquer through contour surgery} \label{sec:surgery}
Now that we have an efficient join tree algorithm, we now describe how to 
reduce computing the contour tree to computing the join tree, as discussed in \Sec{approach}.
Specifically, in this section we first describe our basic cutting tool, which we call contour surgery.
In \Sec{rain} we then describe how we use this tool to cut the surface into so called extremum dominant pieces 
where computing the contour tree reduces to computing the join tree, and this equivalence is proven in \Sec{extreme}. 
The pieces we create are technically not simplicial complexes,
in that each piece may have facets that are not simplices, that is facets of larger size than the dimension.
These can easily be completed into simplicial complexes, though as we do not
need this, for ease of exposition, we do not explicitly
perform this step in our algorithm.
